Case Study: ChIP-seq Differential Peak Calling (csaw)

### 2018-07-05 datasets/ChIP-seq/ChIPseq-CBP-csaw.Rmd
Add random (uninformative) covariate to ChIP-seq

## intermediary files we create below
count_file <- file.path(resdir, "cbp-csaw-counts.rds")
filtered_file <- file.path(resdir, "cbp-csaw-counts-filtered.rds")
normfacs_file <- file.path(resdir, "cbp-csaw-counts-normfacs.rds")
result_file <- file.path(resdir, "cbp-csaw-results.rds")
bench_file <- file.path(sbdir, "cbp-csaw-benchmark.rds")

@@

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)
```

@@

```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
```

@@

@@

@@

>>>>

## intermediary files we create below
count_file <- file.path(resdir, "cbp-csaw-counts.rds")
filtered_file <- file.path(resdir, "cbp-csaw-counts-filtered.rds")
normfacs_file <- file.path(resdir, "cbp-csaw-counts-normfacs.rds")
result_file <- file.path(resdir, "cbp-csaw-results.rds")
bench_file <- file.path(sbdir, "cbp-csaw-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "cbp-csaw-uninf-benchmark.rds")

@@

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)

## add random covariate 
set.seed(11245)
res_df$uninf_covar = rnorm(nrow(res_df))
```

@@

```{r, width-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
```

### Random

```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "uninf_covar") +
    ggtitle("Random independent covariate") +
    xlab("Region Width")
```

```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35, numQ = 3)
```

@@

We'll also compare to the random covariate.

```{r}
if (!file.exists(bench_file_uninf)) {
    res_df$ind_covariate <- res_df$uninf_covar
    res_df$pval <- res_df$PValue

    bd <- initializeBenchDesign()
    sbU <- buildBench(bd, data = res_df, ftCols = "ind_covariate")
    saveRDS(sbU, file = bench_file_uninf)
} else {
    sbU <- readRDS(bench_file_uninf)
}
```

@@

### Region width

@@

### Random

We'll do the same for the random (uninformative covariate).

```{r, chipseq-metrics-rand}
assayNames(sbU) <- "qvalue"
sbU <- addDefaultMetrics(sbU)
```

```{r, chipseq-scatter-rand, results="hide"}
rejections_scatter(sbU, supplementary = FALSE)
rejection_scatter_bins(sbU, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r, chipseq-overlap-rand}
plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r, rnaseq-lineplot-rand}
covariateLinePlot(sbU, alpha = 0.05, covname = "ind_covariate")
```

## Covariate comparison

Here we compare the method ranks for the two covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(benchfile, benchfile_uninf), 
                colLabels = c("Region width", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```

### 2018-07-05 datasets/ChIP-seq/ChIPseq-h3k4me3-csaw.Rmd
Add random (uninformative) covariate to ChIP-seq

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
filtered_file <- file.path(resdir, "h3k4me3-csaw-counts-filtered.rds")
result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
bench_file <- file.path(sbdir, "h3k4me3-csaw-benchmark.rds")

@@

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)
```

@@

```{r, meancov-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
```

@@

>>>>

## intermediary files we create below
count_file <- file.path(resdir, "h3k4me3-csaw-counts.rds")
filtered_file <- file.path(resdir, "h3k4me3-csaw-counts-filtered.rds")
result_file <- file.path(resdir, "h3k4me3-csaw-results.rds")
bench_file <- file.path(sbdir, "h3k4me3-csaw-benchmark.rds")
bench_file_uninf <- file.path(sbdir, "h3k4me3-csaw-uninf-benchmark.rds")

@@

## covert to df for downstream analysis
res_df <- as.data.frame(res_ranges)

## add random covariate 
set.seed(7778)
res_df$uninf_covar = rnorm(nrow(res_df))
```

@@

```{r, width-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "width", maxy = 35, numQ = 3)
```

### Random

```{r, rand-diag-scatter, fig.width=4.5, fig.height=3.5}
rank_scatter(res_df, pvalue = "PValue", covariate = "uninf_covar") +
    ggtitle("Region width as independent covariate") +
    xlab("Region Width")
```

```{r, rand-diag-hist, fig.width=10, fig.height=3.2}
strat_hist(res_df, pvalue = "PValue", covariate = "uninf_covar", maxy = 35, numQ = 3)
```

@@

### Random

We'll do the same for the random (uninformative covariate).

```{r, chipseq-metrics-rand}
assayNames(sbU) <- "qvalue"
sbU <- addDefaultMetrics(sbU)
```

```{r, chipseq-scatter-rand, results="hide"}
rejections_scatter(sbU, supplementary = FALSE)
rejection_scatter_bins(sbU, covariate = "ind_covariate",
                       bins = 4, supplementary = FALSE)
```

```{r, chipseq-overlap-rand}
plotFDRMethodsOverlap(sbU, alpha = 0.05, nsets = ncol(sb),
                      order.by = "freq", decreasing = TRUE,
                      supplementary = FALSE)
```

```{r, rnaseq-lineplots-rand}
covariateLinePlot(sbU, alpha = 0.05, covname = "ind_covariate")
```

## Covariate comparison

Here we compare the method ranks for the two covariates at alpha = 0.10.

```{r}
plotMethodRanks(c(benchfile, benchfile_uninf), 
                colLabels = c("Region width", "Random"), 
                alpha = 0.10, xlab = "Covariate", 
                excludeMethods = NULL)
```
